---
title: "NGS580 Analysis Report"
author: '`r Sys.info()[["user"]] `'
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc: yes
    df_print: kable
  html_document:
    css: styles.css
    df_print: paged
    keep_md: yes
    number_sections: yes
    toc: yes
params:
  old_unpaired_annot: old_unpaired_annot.tsv
  new_unpaired_annot: new_unpaired_annot.tsv
  old_paired_annot: old_paired_annot.tsv
  new_paired_annot: new_paired_annot.tsv
  old_unpaired_annot_path: old_unpaired_annot.tsv
  new_unpaired_annot_path: new_unpaired_annot.tsv
  old_paired_annot_path: old_paired_annot.tsv
  new_paired_annot_path: new_paired_annot.tsv
---
```{r setup, include=FALSE}
# This is the main parent report file, loads and compiles all child documents for the report
library("ggplot2")
library("knitr")
source("overlap.R")

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.height = 12)
knitr::opts_chunk$set(fig.width = 12)

# actual files to be read in
old_unpaired_annot_file <- params$old_unpaired_annot
new_unpaired_annot_file <- params$new_unpaired_annot
old_paired_annot_file <- params$old_paired_annot
new_paired_annot_file <- params$new_paired_annot

# original source paths to the files; Nextflow stages the files so this might not be the same as the original file
old_unpaired_annot_path <- params$old_unpaired_annot_path
new_unpaired_annot_path <- params$new_unpaired_annot_path
old_paired_annot_path <- params$old_paired_annot_path
new_paired_annot_path <- params$new_paired_annot_path



overlap_barplot <- function(overlap_aggr_df, plot_title = "Plot Title"){
    p <- ggplot(data = overlap_aggr_df, 
                aes(x = all, y = pcnt, fill = comb)) +
        geom_bar(stat = "identity", position="stack") +
        coord_flip() +
        theme_bw() +
        ggtitle(plot_title) +
        theme(
            panel.grid.minor.x = element_blank(),
            panel.grid.major.x = element_blank(),
            axis.ticks = element_blank(),
            axis.text.x = element_blank(),
            axis.title.x = element_blank()
        )
    return(p)
}
```

```{r load_data}
old_unpaired_annot <- read.delim(file = old_unpaired_annot_file, 
                                  header = TRUE, 
                                  sep = '\t', 
                                  check.names = FALSE)
new_unpaired_annot <- read.delim(file = new_unpaired_annot_file, 
                                  header = TRUE, 
                                  sep = '\t', 
                                  check.names = FALSE)
old_paired_annot <- read.delim(file = old_paired_annot_file, 
                                  header = TRUE, 
                                  sep = '\t', 
                                  check.names = FALSE)
new_paired_annot <- read.delim(file = new_paired_annot_file, 
                                  header = TRUE, 
                                  sep = '\t', 
                                  check.names = FALSE)

save.image(file="loaded.Rdata",compress = TRUE)
```

```{r separate}
# separate based on variant callers
unpaired_variants <- list()
# get unique set of callers and caller types
unpaired_caller_types <- rbind(
    old_unpaired_annot[, c("VariantCaller", "VariantCallerType")],
    new_unpaired_annot[, c("VariantCaller", "VariantCallerType")]
)
unpaired_caller_types <- unpaired_caller_types[!duplicated(unpaired_caller_types),]

for(i in seq(nrow(unpaired_caller_types))){
    vc <- as.character(unpaired_caller_types[i, "VariantCaller"])
    vct <- as.character(unpaired_caller_types[i, "VariantCallerType"])
    
    unpaired_variants[[vc]] <- list()
    unpaired_variants[[vc]][[vct]] <- list()
    
    # filter out for only the matching variants from each dataset
    unpaired_variants[[vc]][[vct]][["old"]] <- old_unpaired_annot[
        which(old_unpaired_annot[["VariantCaller"]] == vc & old_unpaired_annot[["VariantCallerType"]] == vct ) ,]
    # create unique variant ID for each entry
    unpaired_variants[[vc]][[vct]][["old"]][['VariantID']] <- sprintf(
        "%s:%s:%s>%s:%s", 
        unpaired_variants[[vc]][[vct]][["old"]][['CHROM']], 
        unpaired_variants[[vc]][[vct]][["old"]][['POS']], 
        unpaired_variants[[vc]][[vct]][["old"]][['REF']], 
        unpaired_variants[[vc]][[vct]][["old"]][['ALT']],
        unpaired_variants[[vc]][[vct]][["old"]][['Sample']]
    )
    
    unpaired_variants[[vc]][[vct]][["new"]] <- new_unpaired_annot[
        which(new_unpaired_annot[["VariantCaller"]] == vc & new_unpaired_annot[["VariantCallerType"]] == vct ) ,]
    unpaired_variants[[vc]][[vct]][["new"]][['VariantID']] <- sprintf(
        "%s:%s:%s>%s:%s", 
        unpaired_variants[[vc]][[vct]][["new"]][['CHROM']], 
        unpaired_variants[[vc]][[vct]][["new"]][['POS']], 
        unpaired_variants[[vc]][[vct]][["new"]][['REF']], 
        unpaired_variants[[vc]][[vct]][["new"]][['ALT']],
        unpaired_variants[[vc]][[vct]][["new"]][['Sample']]
    )
    
    # create the overlap table
    unpaired_variants[[vc]][[vct]][["overlap"]] <- create_overlaps(
        variant_list = list(
            new = unpaired_variants[[vc]][[vct]][["new"]][['VariantID']],
            old = unpaired_variants[[vc]][[vct]][["old"]][['VariantID']]
        )
    )
    
    # create aggregate table
    unpaired_variants[[vc]][[vct]][["overlap_aggr"]] <- aggregate_overlaps(unpaired_variants[[vc]][[vct]][["overlap"]])
    
    # make plot
    unpaired_variants[[vc]][[vct]][["overlap_aggr_plot"]] <- overlap_barplot(unpaired_variants[[vc]][[vct]][["overlap_aggr"]])
}

paired_variants <- list()
paired_caller_types <- rbind(
    old_paired_annot[, c("VariantCaller", "VariantCallerType")],
    new_paired_annot[, c("VariantCaller", "VariantCallerType")]
)
paired_caller_types <- paired_caller_types[!duplicated(paired_caller_types),]

for(i in seq(nrow(paired_caller_types))){
    vc <- as.character(paired_caller_types[i, "VariantCaller"])
    vct <- as.character(paired_caller_types[i, "VariantCallerType"])
    paired_variants[[vc]] <- list()
    paired_variants[[vc]][[vct]] <- list()
    paired_variants[[vc]][[vct]][["old"]] <- old_paired_annot[
        which(old_paired_annot[["VariantCaller"]] == vc & old_paired_annot[["VariantCallerType"]] == vct )
        ,]
    paired_variants[[vc]][[vct]][["new"]] <- new_paired_annot[
        which(new_paired_annot[["VariantCaller"]] == vc & new_paired_annot[["VariantCallerType"]] == vct )
        ,]
    
    paired_variants[[vc]][[vct]][["old"]][['VariantID']] <- sprintf(
        "%s:%s:%s:%s:%s:%s", 
        paired_variants[[vc]][[vct]][["old"]][['CHROM']], 
        paired_variants[[vc]][[vct]][["old"]][['POS']], 
        paired_variants[[vc]][[vct]][["old"]][['REF']], 
        paired_variants[[vc]][[vct]][["old"]][['ALT']],
        paired_variants[[vc]][[vct]][["old"]][['Tumor']],
        paired_variants[[vc]][[vct]][["old"]][['Normal']]
    )
    paired_variants[[vc]][[vct]][["new"]][['VariantID']] <- sprintf(
        "%s:%s:%s:%s:%s:%s", 
        paired_variants[[vc]][[vct]][["new"]][['CHROM']], 
        paired_variants[[vc]][[vct]][["new"]][['POS']], 
        paired_variants[[vc]][[vct]][["new"]][['REF']], 
        paired_variants[[vc]][[vct]][["new"]][['ALT']],
        paired_variants[[vc]][[vct]][["new"]][['Tumor']],
        paired_variants[[vc]][[vct]][["new"]][['Normal']]
    )
    
    paired_variants[[vc]][[vct]][["overlap"]] <- create_overlaps(
        variant_list = list(
            new = paired_variants[[vc]][[vct]][["new"]][['VariantID']],
            old = paired_variants[[vc]][[vct]][["old"]][['VariantID']]
        )
    )
    
    paired_variants[[vc]][[vct]][["overlap_aggr"]] <- aggregate_overlaps(paired_variants[[vc]][[vct]][["overlap"]])
    
    # make plot
    paired_variants[[vc]][[vct]][["overlap_aggr_plot"]] <- overlap_barplot(paired_variants[[vc]][[vct]][["overlap_aggr"]])
}
save.image(file="separated.Rdata",compress = TRUE)
```

```{r metrics}
annot_metrics_unpaired <- function(df){
    df[["n"]] <- 1
    df <- aggregate(n ~ VariantCaller + VariantCallerType + Sample, data = df, FUN = sum)
    return(df)
}
annot_metrics_paired <- function(df){
    df[["n"]] <- 1
    df <- aggregate(n ~ VariantCaller + VariantCallerType + Tumor + Normal, data = df, FUN = sum)
    return(df)
}
metrics <- list()
metrics[["unpaired"]] <- list()
metrics[["unpaired"]][["old"]] <- annot_metrics_unpaired(old_unpaired_annot)
metrics[["unpaired"]][["new"]] <- annot_metrics_unpaired(new_unpaired_annot)
metrics[["paired"]] <- list()
metrics[["paired"]][["old"]] <- annot_metrics_paired(old_paired_annot)
metrics[["paired"]][["new"]] <- annot_metrics_paired(new_paired_annot)

```

# Metrics

## Old Unpaired Annotations

```{r}
metrics[["unpaired"]][["old"]]
```

- `r old_unpaired_annot_path`


## New Unpaired Annotations

```{r}
metrics[["unpaired"]][["new"]]
```

- `r new_unpaired_annot_path`

## Old Paired Annotations

```{r}
metrics[["paired"]][["old"]]
```

- `r old_paired_annot_path`

## New Paired Annotations

```{r}
metrics[["paired"]][["new"]]
```
- `r new_paired_annot_path`

# Annotation Overlap

## Overall Overlap

Comparison of all variant annotations between old and new annotation lists

# Unpaired Variants

```{r results='asis'}
for (vc in names(unpaired_variants)){
    for(vct in names(unpaired_variants[[vc]])){
        cat(sprintf("## %s.%s\n\n", vc, vct))
        # print(unpaired_variants[[vc]][[vct]][["overlap_aggr"]])
        print(unpaired_variants[[vc]][[vct]][["overlap_aggr_plot"]])
        cat("\n\n")
    }
}
```

# Paired Variants

```{r results='asis'}
for (vc in names(paired_variants)){
    for(vct in names(paired_variants[[vc]])){
        cat(sprintf("## %s.%s\n\n", vc, vct))
        # print(paired_variants[[vc]][[vct]][["overlap_aggr"]])
        print(paired_variants[[vc]][[vct]][["overlap_aggr_plot"]])
        cat("\n\n")
    }
}
```

# System Information 

```{r sysinfo}
sessionInfo()
save.image(file="final.Rdata",compress = TRUE)
```
